---
title: "Sports Analytics"
author: "Usman Derek Yahaya"
date: "2023-03-06"
output: html_document
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

I had analyzed the 2022 fifa world cup, using StatsBomb i extracted and analysed the shots and goals totals for each team, then explored average shots and goals per game.  

## Install Libraries
```{r}
library(ggplot2)
library(StatsBombR)
```

## Filter Competition and Season
This filters the competitions to return the world cup data, using the dplyrʼs ʻfilterʼ function, to just the 2022 FIFA World Cup. 
```{r, echo=FALSE}
Comp <- FreeCompetitions() %>%
  filter(competition_id==43 & season_name=="2022")
```

## Extract Matches Data and Assisgn to Dataframe
```{r}
Matches <- FreeMatches(Comp)
```

## Assign StatsBombData event data for 2022 FIFA World Cup
```{r}
StatsBombData <- free_allevents(MatchesDF = Matches, Parallel = T)
```

## Clean Dataframe
```{r}
StatsBombData = allclean(StatsBombData) 
```

## Extract shots and goals totals 
```{r}
shots_goals = StatsBombData %>%
  group_by(team.name) %>%
  summarise(shots = sum(type.name=="Shot", na.rm = TRUE),
            goals = sum(shot.outcome.name=="Goal", na.rm = TRUE))
```

## Display Data Structure
```{r}
str(shots_goals)
```


## Plot Total Shots by Teams 
```{r}
ggplot(data = shots_goals,
       aes(x = reorder(team.name, shots), y = shots)) +
  geom_bar(stat = "identity", width = 0.7, fill = "Red") +
  labs(y="Shots") +
  theme(axis.title.y = element_blank()) +
  scale_y_continuous( expand = c(0,0)) +  
  coord_flip()
```


## Plot Total Goals by Teams 
```{r}
ggplot(data = shots_goals,
       aes(x = reorder(team.name, shots), y = goals)) +
  geom_bar(stat = "identity", width = 0.7, fill = "#00abff") +
  labs(y="Goals") +
  theme(axis.title.y = element_blank()) +
  scale_y_continuous( expand = c(0,0)) +  
  coord_flip()
```
